Teaching Low-Level Concurrency with \pebbles and \landslide

### 2016-08-16 sigcse-17/intro.tex
sigcse: some revisions

In this paper we describe our experience teaching concurrent programming at the operating-system level to advanced undergraduate students at \carnegiemellon (\cmu).
In \fourten, the Operating Systems Design and Implementation class at \cmu,
students complete aggressively low-level thread library and kernel projects,
which run on real hardware,
in which they implement threads and processes.
Presented with a kernel specification called \pebbles%
% TODO camready
%\cite{kspec}
, students build a threading and synchronization library that runs in user space on top of a \pebbles-compliant kernel in a two-week project, and subsequently, in a six-week project, write their own multi-threaded kernel implementation.
During these projects students face demanding synchronization problems, including how to implement locking primitives and how to coordinate the possibly-simultaneous exiting of parent and child processes.
% Finally, in another two-week project, students implement an extension to their Pebbles kernel. This project changes every semester, and recently has involved supporting symmetric multiprocessing (SMP).

>>>>

In this paper we describe our experience teaching concurrent programming at the operating-system level to advanced undergraduate students at \carnegiemellon (\cmu).
In \fourten, the Operating Systems Design and Implementation class at \cmu,
students complete aggressively low-level thread library and kernel projects,
which run on real hardware,
in which they implement threads and processes.
Presented with a kernel specification called \pebbles%
% TODO camready
%\cite{kspec}
, students build a user-space threading and synchronization library on top of a \pebbles-compliant kernel in a two-week project, and subsequently, in a six-week project, implement their own multi-threaded kernel.
%During these projects
The students face demanding synchronization problems, including how to implement locking primitives and how to coordinate the possibly-simultaneous exiting of parent and child processes.
% Finally, in another two-week project, students implement an extension to their Pebbles kernel. This project changes every semester, and recently has involved supporting symmetric multiprocessing (SMP).
